### 2023-01-15 writing.qmd

# Writing

Tutorial and course notes -- shared for reading group and teaching purposes.
Also blog posts to support and explain research papers.

## Notes
### [Probability primer]()
Notes for the ROSEHIPS reading group (TBC).

### [Introduction to Engineering Pattern Recognition](https://github.com/labull/EPR-lectures)
Guest Lectures for the Advanced Dynamics course at the University of Sheffield in the Engineering dept.
A light introduction to machine learning in the context of some engineering examples:

- EPR1 | Classifying acoustic emission signals

- EPR2 | Regression analysis of wind turbine power curves

## Paper summaries