### 2023-01-16 index.qmd
edit index, writing

---
title: "Lawrence Bull, PhD"
about:
  template: solana
  image: images/mugshot.png
  links:
    - text: github
      href: https://github.com/labull
    - text: google scholar
      href: https://scholar.google.com/citations?user=Lszt1B4AAAAJ&hl=en
    - text: research gate
      href: https://www.researchgate.net/profile/Lawrence-Bull-2
    - text: orcid
      href: https://orcid.org/0000-0002-0225-5010
    - text: twitter
      href: https://twitter.com/lwrncebull?lang=ar
---

<div style="text-align: left">

**[2022-2027] University of Cambridge** | [Computational Statistics and Machine Learning Group](https://csml-cam.github.io) | Research Associate | Department of Engineering

**[2021-2022] The Alan Turing Institute** | Research Associate | [Data-Centric Engineering](https://www.turing.ac.uk/research/research-programmes/data-centric-engineering)

**[2016-2021] University of Sheffield** | PhD Student (2016-2019) Research Assistant (2019-2022) | [Dynamics Research Group](https://drg.ac.uk/) | Department of Engineering

</br>

Broadly speaking, I'm a researcher in statistical methods for Data-Centric Engineering. Focussing on monitoring techniques for telemetry data from systems and infrastructure.

I'm particularly interested in _multilevel structures_ that appear in measurements from populations of systems -- trains, [planes](), [automobiles](https://doi.org/10.48550/arXiv.2204.12404) (... [wind turbines](https://arxiv.org/abs/2111.15496)).
By encoding domain expertise in a Bayesian workflow, complex models can be built from interpretable units, allowing machines to share information and increasing the value of data.
At the population level, alternative insights can emerge that might otherwise be missed -- the whole is something _besides_ its parts.

I'm a communications editor for the [Data-Centric Engineering](https://www.cambridge.org/core/journals/data-centric-engineering) journal (Cambridge University Press) and a member of the [Early-Career Academic and Professionals Panel](https://www-smartinfrastructure.eng.cam.ac.uk/news/csic-call-applicants-early-career-academics-and-professionals-panel) (ECAPP) in the [Cambridge Centre for Smart Infrastructure and Construction](https://www-smartinfrastructure.eng.cam.ac.uk/) (CSIC).