### 2024-04-02 index.qmd
glasgow

---
title: "Lawrence Bull, PhD"
about:
  template: solana
  image: images/mugshot.png
---

</br>

<div style="text-align: left">

**[May.2024-May.2027] University of Glasgow** | Research Fellow | [Statistics & Data Analytics](https://www.gla.ac.uk/schools/mathematicsstatistics/research/stats/ai3/) | School of Mathematics and Statistics

**[Sep.2022-May.2024] University of Cambridge** | Research Associate | [Computational Statistics and Machine Learning Group](https://csml-cam.github.io) | Department of Engineering

**[Jul.2021-Sep.2022] The Alan Turing Institute** | Research Associate | [Data-Centric Engineering](https://www.turing.ac.uk/research/research-programmes/data-centric-engineering)

**[Sep.2016-Jun.2021] University of Sheffield** | PhD Student (2016-2019) Research Assistant (2019-2021) | [Dynamics Research Group](https://drg.ac.uk/) | Department of Engineering

</br>

Researcher in statistical methods for Data-Centric Engineering, specialising in monitoring techniques for telemetry data from systems and infrastructure.

I'm interested in the dependencies that appear in measurements from populations of machines -- [planes](https://www.sciencedirect.com/science/article/pii/S0022460X21001449), [trains](https://arxiv.org/abs/2403.17820), [automobiles](https://doi.org/10.48550/arXiv.2204.12404) (... [wind turbines](https://arxiv.org/abs/2111.15496)).
By encoding domain expertise into statistical models, complex representations can be built from interpretable units, allowing machines to share information and increase the value of data.
<!-- At a systems level, alternative insights can emerge that might otherwise be missed -- the whole is something _besides_ its parts. -->

I'm a communications editor for the [Data-Centric Engineering](https://www.cambridge.org/core/journals/data-centric-engineering) journal and a Visiting Research Fellow at [Cambridge University](https://csml-cam.github.io). I was on the [Early-Career Academic and Professionals Panel](https://www-smartinfrastructure.eng.cam.ac.uk/news/csic-call-applicants-early-career-academics-and-professionals-panel) (2021-2023) in the [Cambridge Centre for Smart Infrastructure and Construction](https://www-smartinfrastructure.eng.cam.ac.uk/) (CSIC).

<div style="text-align: right">

</br>

[github](https://github.com/labull)

[google scholar](https://scholar.google.com/citations?user=Lszt1B4AAAAJ&hl=en)

[research gate](https://www.researchgate.net/profile/Lawrence-Bull-2)

[orcid](https://orcid.org/0000-0002-0225-5010)

[twitter](https://twitter.com/lwrncebull?lang=ar)