### 2024-04-24 writing.qmd

# Writing

Tutorial and course notes -- shared for reading group and teaching purposes.
Also blog posts to support and explain research papers.

## Notes

### [Introduction to Engineering Pattern Recognition](https://github.com/labull/EPR-lectures)
Initially guest Lectures for the Advanced Dynamics [MEC452] undergraduate course at the University of Sheffield in the Engineering Dept.

Expanded to a workshop for the [IMAC conference, 2024](https://sem.org/pre-con-course-2).

A light introduction to machine learning in the context of some engineering examples:

- EPR1 | Classifying acoustic emission signals

- EPR2 | Regression analysis of wind turbine power curves

## Paper summaries